Networks Matter...or some nonsense like that

### 2016-11-25 amenIntro/amenIntro.Rmd
init review done

* Why are we meeting during Thanksgiving weekend...I didn't realize the date and none of you objected :)
* Anyhow...we are here to talk about conducting a meta-analysis of dyadic studies in political science using the Additive and Multiplicative Effects Model for Networks (AMEN)
* What I hope we find from this project is that using standard GLM approaches on dyadic data will produce
      - biased estimates of the effect of independent variables
      - uncalibrated confidence intervals
      - and poor predictive performance.
* Anticipated outputs: 
      - One big meta-analysis piece in which we show the items in the list above to be true on average
      - A number of small replication pieces

@@

## Assessing convergence

## Did you set the right K?

## Exogenous covariates

```{r, echo=TRUE}
dim(fit$BETA)

stats = function(x){ c( mu=mean(x), q95  ) }
apply(BETA<)
```

@@

>>>>

* Why are we meeting during Thanksgiving weekend...I didn't realize the date and none of you objected :)
* Anyhow...we are here to talk about conducting a meta-analysis of dyadic studies in political science using the Additive and Multiplicative Effects Model for Networks (AMEN)
* What I hope we find from this project is that using standard GLM approaches on dyadic data produces
      - biased estimates of the effect of independent variables
      - uncalibrated confidence intervals
      - and poor predictive performance.
* Anticipated outputs: 
      - One big meta-analysis piece in which we show the items in the list above to be true on average
      - A number of small replication pieces
      - New disciples to spread the gospel

@@

## First step...did the model converge

```{r, echo=TRUE}
fit$BETA[1:3,]
```

```{r, echo=TRUE, eval=FALSE}
library(reshape2) ; library(ggplot2)

beta = data.frame(fit$BETA) ; names(beta)[ncol(beta)] = 'distance'
beta$iter = 1:nrow(beta)*10

ggBeta = reshape2::melt(beta,id='iter')
ggplot(ggBeta, aes(x=iter,y=value)) + 
	geom_line() + 
	facet_wrap(~variable,nrow=2,scales='free_y')
```

## First step...did the model converge

```{r, echo=FALSE}
library(reshape2) ; library(ggplot2)
beta = data.frame(fit$BETA) ; names(beta)[ncol(beta)] = 'distance'
beta$iter = 1:nrow(beta)*10
ggBeta = reshape2::melt(beta,id='iter')
ggplot(ggBeta, aes(x=iter,y=value)) + geom_line() + facet_wrap(~variable,nrow=2,scales='free_y')
```

## Are we capturing the dependencies in our data?

* Lets find out by drawing from the posterior predictive distribution of our model

```{r, echo=TRUE}
fit$GOF[1:3,]
```

```{r,message=FALSE, warning=FALSE, suppressPackageStartupMessages=TRUE,fig.width=8, fig.height=2}
act = fit$GOF[1,-c(2:3)] ; gof = fit$GOF[-1,-c(2:3)]

ggGof = melt(gof)[,-1]
ggGof$act = act[match(ggGof$Var2, names(act))]

ggplot(ggGof, aes(x=value)) + geom_histogram() + 
	facet_wrap(~Var2,nrow=1,scales='free') + 
	geom_vline(aes(xintercept=act),color='red')
```

## Parameter estimates

```{r, echo=TRUE}
library(magrittr) # so it's easier to see
summStats = function(x){ c(mu=mean(x), sd=sd(x), quantile(x, probs=c(0.025,0.975)))  }

fit$BETA %>% apply(., 2, summStats) %>% round(., 3) %>% t(.)
```
* You would then compare these with the corresponding GLM model.

## Visualizing third order dependence

```{r, echo=TRUE, fig.height=5, fig.width=5, fig.align='center'}
circplot(Y=yL[[length(yL)]], U=fit$U, lcol='white')
```

@@

* Randomly divide the $n \times (n-1)$ data points into $S$ sets of roughly equal size, letting $s_{ij}$ be the set to which pair $\{ij\}$ is assigned.
* For each $s \in \{1, \ldots, S\}$:
	- Obtain estimates of the model parameters conditional on $\{y_{ij} : s_{ij} \neq s\}$, the data on pairs not in set $s$.
	- For pairs $\{kl\}$ in set $s$, let $\hat y_{kl} = E[y_{kl} | \{y_{ij} : s_{ij} \neq s\}]$, the predicted value of $y_{kl}$ obtained using data not in set $s$.

* Code to perform this analysis for the example is shown in `git/netsMatter/code/example/runOutPerf.R`
* When we get out of sample results we will be evaluating across two metrics:
	- Receiver Operating Characteristic curves
	- Precision-Recall curves

## There will be a lot of models to run...

```{r, eval=FALSE, echo=TRUE}
# run ame in parallel # mcmc params
imps = 100000
brn = 50000
ods = 10
latDims = 1:4
seed=6886

# Run amen in parallel
library(doParallel) ; library(foreach)
cl=makeCluster(4) ; registerDoParallel(cl)
foreach(ii=1:length(latDims), .packages=c("amen")) %dopar% {
	ameFit = ame_repL(
		Y=yL,Xdyad=xDyadL,Xrow=xNodeL,Xcol=NULL, model="bin",symmetric=TRUE,
		R=latDims[ii], 
		nscan=imps, seed=seed, burn=brn, odens=ods, 
		plot=FALSE, print=FALSE) 	
	save(ameFit, 
		file='~/Dropbox/netsMatter/replications/example/outputData/model_k',
			latDims[ii],'.rda')
}
stopCluster(cl)
```

## Code + Data + Conversation

* We will be writing code on Git **ONLY**
* Input and output data will be stored on Dropbox **ONLY**
* We can converse on Slack 
	- As you guys run into issues/questions/concerns we can raise them there and answer them in close to real-time

## Project Management

* Teams of two, each of you will be paired with someone experienced in AME
	- Dorff + Foster
	- Gallop + Liu
	- Ward + Roberts
	- Me + Tellez 
* Claim the replications you want to do on Google sheets ... whoever gets there first wins.
	- Mansfield + Milner is already mine.
* <span style="color:red">**How many replications should we do?**</span>
* <span style="color:red">**Timeline?**</span>
* <span style="color:red">**What else?**</span>
